---
output: html_document
---
#Reproducible Researh - Assigment 1
##Juan Guillermo López Guzmán

In this document we are working a database based on anonymous users data. You can download the database in this url: [database](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

```{r loading database,echo=FALSE}
        db<-read.csv("C:/Users/FAMILIAR/Documents/activity.csv")
```

This is a briefly summary of database
```{r Summary}
        summary(db)
```

##What is mean total number of steps taken per day?  
Firstly we will get the total steps in each day for that we will use summarize function from the dplyr library
```{r Total steps per day,message=FALSE}
        library(dplyr)
        summarize(group_by(db,date),sum=sum(db$steps))
```
To analize the distribution of the total steps in days we will make an histogram:
```{r,echo=FALSE}
        plotdb<-summarize(group_by(db,date),steps=sum(db$steps))
        hist(plotdb$steps,main="Total steps per day",xlab="")
```
```{r,echo=FALSE,results='hide'}
        mean<-as.integer(mean(plotdb$steps,na.rm=TRUE))
        median<-as.integer(median(plotdb$steps,na.rm=TRUE))
````
In accordance with the summary we can say the steps mean per day is `r mean` and the steps median per day is `r median`

##What is the average daily activity pattern?  
A time series plot will show the pattern across the day:
```{r activity pattern,echo=FALSE}
        plotdb<-summarize(group_by(db,interval),steps=mean(db$steps,na.rm=TRUE))
        plot(plotdb,xlab="5-minutes interval",ylab="Average steps per interval",type="l")
```

Now we need calculate the 5-minutes intervale with the maximum number of steps average. Firstly identify the maximum number after that we find the interval who contained that average:
```{r max steps interval}
        max_steps<-max(plotdb$steps)
        filter<-plotdb$steps==max_steps
        max_interval<-plotdb$interval[filter]
```
So now we know the maximum number of average steps is `r max_steps` in the `r max_interval` 5-minutes interval

##Imputing missing values

The database have a 2304 total of missing value therefore to improve the data quality we will replace the NA's value with the average steps in the 5-minutes interval:
```{r replacing NA values}
        imp_db<-db
        for(i in 1:length(imp_db$steps)){
                int<-imp_db$interval[i]
                if(is.na(imp_db$steps[i])){
                        imp_db$steps[i]<-mean(filter(db,interval==int)$steps,na.rm=TRUE)
                }
        }
```
```{r,results='hide',echo=FALSE}
        mean<-as.integer(mean(summarize(group_by(imp_db,date),steps=sum(imp_db$steps))$steps,na.rm=TRUE))
        median<-as.integer(median(summarize(group_by(imp_db,date),steps=sum(imp_db$steps))$steps,na.rm=TRUE))
```
The new database have an `r mean` average steps per day and `r median` median. We will see the new distribution with a histogram:
```{r histogram in improved database}
        plotdb<-summarize(group_by(imp_db,date),steps=sum(imp_db$steps))
        hist(plotdb$steps,main="Total steps per day",xlab="")
```

##Are there differences in activity patterns between weekdays and weekends?

To examine this question we will use the `weekdays()` function and add a column with the factor weekend or weekdays
```{r}
        imp_db$date<-strptime(imp_db$date,format="%Y-%m-%d")
        imp_db$WD<-weekdays(imp_db$date)
        for(i in 1:length(imp_db$WD)){
                if(imp_db$WD[i]=="sábado"||imp_db$WD[i]=="domingo"){
                        imp_db$WD[i]<-"weekend"
                        }else{
                                imp_db$WD[i]<-"weekday"
                                }
                }
```
In adition we plot a comparative timeline graph between weekdays and weekends:
```{r,echo=FALSE}
        imp_db<-imp_db[-2]        
        plotdb<-summarize(group_by(imp_db,WD,interval),steps=mean(imp_db$steps,na.rm=TRUE))
        library(lattice)
        xyplot(steps~interval|WD,data=plotdb,type="l",layout=c(1,2))
        wday_mean<-mean(filter(imp_db,WD=="weekday")$steps)
        wend_mean<-mean(filter(imp_db,WD=="weekend")$steps)
```

This graphic shows how the users walk more on weekend with an average steps per interval of `r wend_mean` versus the `r wday_mean` in weekdays